---
title: "FMR_Raw_Analysis"
output: html_document
---

###  Analysis of metabolic rate, somewhat taken from Michael Dillon and Abbie Reade (MDAR) scripts...
# for analysis in R of Expedata files previously exported as .csv 
#  
# These are for Stephen's honeybee experiments, flying chamber bees, 350 mL chamber, with channels:
#
#    
#    Seconds: seconds, starts with "0"
#    Barometric_Pressure:  pressure (kPa)
#    Flow_Rate: flow rate in system (mL/min)
#    Oxygen:   Oxygen %, corrected for barometric pressure
#    Water_vapor:   Water vapor (kPa) *******(mmol/mol so divide by 1000 to get molar fraction)
#    Carbon_Dioxide:   CO2 %, corrected for barometric pressure

```{r}
library(dplyr)
library(stringr)
library(foreach)
library(segmented)
library(grDevices)
library(tidyr)
library(zoo)
library(tidyverse)

#set the working directory, here, it is reinforcing the default project directory
setwd("C:/Users/jcassano/Desktop/Masters Thesis/Summer 2020 Thesis Project/FMR Analysis/FMR Raw Data")

# directory for visual checks of data - IDK what this is!
dir.checks <- "output/checktrace"
# directory where all corrected data is housed for further analysis
dir.processed <- "output/processed"
```

#Reading in data
```{r}
# Functions to load in expedata files saves as CSV
load.expedata.file <- function(){
  file <- file.choose()
  # extract information from filename
  filename <<- unlist(strsplit(basename(file), "[ ]"))[1]
  filename <<- gsub(".csv", "", tail(strsplit(file, "/")[[1]], n=1))
  #edate    <<- strsplit(filename, "_")[[1]][2]
  # read in data to "flightfull" which will be full resting data
  flightfull   <- read.csv(file, header=T)
  # Markers .. these are saved as Ascii codes, so here we convert them back to characters
  flightfull$markers <- rawToChar(as.raw(as.character(flightfull$Marker)), multiple=T)
  # this returns decimal fraction, uncorrected for water dilution because lags have to be fixed first
  #flightfull$CO2uc <- flightfull$CO2/1000000 
  unique(flightfull$chamber)
  return(flightfull)
  #write.csv(flightfull, "test.csv", row.names=F)
}

#Use function to load CSV files
#Choose files from directory
flightfull <- load.expedata.file()%>%
  dplyr::rename("Carbon_Dioxide" = "CO2",
         "Oxygen" = "O2")%>%
  dplyr::select(-c("Temp__Aux1_","Temp__Aux2_", "FoxBoxTemp", "FlowSet")) #select variables of interest
str(flightfull)
View(flightfull)
```

```{r}
#create new marker names TO CHECK WORK
#flightfull$Markname <- ifelse(flightfull$Marker == -1, flightfull$Markname <- "none", #if Marker = 1, mark name 
                        #ifelse(flightfull$Marker == 108, flightfull$Markname <- "end", 
                       # ifelse(flightfull$Marker == 118, flightfull$Markname <- "beginbase", flightfull$Markname <- "check")))

#############
#flightfull$Markname <- rep(NA, 804)
#for(i in flightfull$Seconds){
  #if(flightfull[i+1,6]==-1){flightfull[i+1,8] <- "none"}
  #else if(flightfull[i+1,6] == 108) {flightfull[i+1,8] <- "end"}
  #else if(flightfull[i+1,6] == 118) {flightfull[i+1,8] <- "beginbase"}
  #else (flightfull[i+1,8] <- "check")
#}
```

```{r}
#Timepoint determination
#This code populates the blanks in markers with NA's then fills in every NA with a previous actual value entry, effectively making each row fall into a timepoint position based on the markers column
flightfull$markers[1]<-"K" #make first value "K" signifying the beginning
flightfull$markers[flightfull$markers == ""] <- NA #populates blank cells with "NA"
flightfull$markers <- na.locf(flightfull$markers) #replaces NAs with repeats of the previous letters
flightfull$markers <- toupper(flightfull$markers) #replaces lower case letters with uppercase letters
```

```{r}
# saves for later use
#This sorts the rows with certain letters and saves them as objects
seal <- which(flightfull$markers == "V") # seal, continues until 1st behavior noted 
fly <- which(flightfull$markers == "F") #flying
bsln <- which(flightfull$markers == "S") #Baseline at beginning
rest <- which(flightfull$markers == "R") #rest

#saves the first row number of each section
seal1 <- seal[1]
fly1 <- fly[1] #flying
bsln1 <- bsln[1] #Baseline at beginning
rest1 <- rest[1] #rest
```

LAG CORRECT
```{r}
#Plot lag correct WHAT IS THIS DOING
plot(flightfull$Seconds[(seal1-20):(seal1+20)], flightfull$Carbon_Dioxide[(seal1-20):(seal1+20)])
abline(v=seal1)
```

```{r}
#This creates a new column with "shifted" or lag-corrected readings to match the time point
lagpres <- 10
lagflow <- 10
#lagwvp <- 10
lago2 <- 10
lagco2 <- 10

flightfull$co2nolag <- c(flightfull$Carbon_Dioxide[(1+lagco2):length(flightfull$Carbon_Dioxide)], rep(NA, lagco2))
flightfull$o2nolag <- c(flightfull$Oxygen[(1+lago2):length(flightfull$Oxygen)], rep(NA, lago2))
flightfull$flownolag <- c(flightfull$Flow_Rate[(1+lago2):length(flightfull$Flow_Rate)], rep(NA, lagflow))
flightfull$presnolag <- c(flightfull$Barometric_Pressure[(1+lagpres):length(flightfull$Barometric_Pressure)], rep(NA, lagpres))
```

```{r}
#plot full data lag corrected
  #plot(flightfull$Seconds, flightfull$wvpnolag, type="l", col="blue")
  #par(new=T)
plot(flightfull$Seconds, flightfull$co2nolag, type="l", col="darkorange", axes=T, ann=F)
par(new=T)
plot(flightfull$Seconds, flightfull$o2nolag, type="l", col="green", axes=T, ann=F)
par(new=T)
plot(flightfull$Seconds, flightfull$flownolag, type="l", axes=T, ann=F)
abline(v=fly1-10)

plot(flightfull$Seconds, flightfull$co2nolag, type="l", col="darkorange", axes=T, ann=F)
```

```{r}
## Drift correct and zero CO2 ----
flightfull$co2_wvpc <- flightfull$co2nolag
allVarBase <- data.frame(startTime=integer(0), varValBase=numeric(0))

for(x in bsln1:(seal1-5)){ 
  varNowBase <- var(flightfull$co2_wvpc[x:(x+30)], na.rm = T)
  newRowBase <- c(x, varNowBase)
  allVarBase <- rbind(allVarBase, newRowBase)  
}

names(allVarBase)=c("startTime", "varValBase")
#View(allVar)
#now to find minumum variance point using the above time frame in the for loop (2 minutes) 
minVarBase <- min(allVarBase$varValBase)

#Get the estimates 
begin.base <- allVarBase$startTime[which(allVarBase$varValBase==min(allVarBase$varValBase))]
duration.base <- 30
lowVar.baseline <- begin.base:(begin.base+duration.base)
prelowvar <- mean(flightfull$co2_wvpc[lowVar.baseline], na.rm=T)

# drift correction
shiftedco2 <- flightfull$co2_wvpc - prelowvar
# subtracting out pre-baseline ("zeroing") 

flightfull$co2pct <- shiftedco2 
```

```{r}
plot(flightfull$Seconds, flightfull$co2pct, type="l")
abline(v=fly1)
plot(flightfull$Seconds[fly], flightfull$co2pct[fly])
abline(v=fly1)
```

```{r}
## Add VCO2 to data frame and save complete corrected trace to file ----
flightfull$VCO2  <- flightfull$co2pct*flightfull$Flow_Rate*60 # mL CO2/hr

plot(flightfull$Seconds[fly], flightfull$VCO2[fly], main=c(filename, "VCO2"), type="l", xlab = "Seconds flight", ylab = "VCO2 flight")
```


```{r}
## Estimate metabolic rate for specified regions and plot for confirmation

# Get rid of the "cap" points that may interrupt continuous flight

##### Subset just flight points -----

# This section creates a new column telling whether the datapoint starts a new behavior, ends it, or is the same as previous
wholetrace <- flightfull$Seconds[1]:length(flightfull$Seconds)
flightfull$markz <- "problem"
for(x in wholetrace) {
ifelse (flightfull$markers[x] != flightfull$markers[x-1], flightfull$markz[x] <- "new", 
                            ifelse (flightfull$markers[x] != flightfull$markers[x+1], flightfull$markz[x] <- "end", flightfull$markz[x] <- "same"))
}
flightfull$markz[length(flightfull$Seconds)] <- "end"
news <- which(flightfull$markz=="new")
for(x in news) {
if(flightfull$markz[x+1] == "new") {flightfull$markz[x] <- "weirdo"}
}
flightfull$Seconds[which(flightfull$markz == "weirdo")]
```
```{r}
# All from 1st flight
#startflight <- data.frame(subset(flightfull, flightfull$Seconds > fly1))

# Selects non-flight
#notflight <- data.frame(subset(flightfull, flightfull$markers != "F"))
```

```{r}
# Selects all "flight bouts"
justflight <- as_tibble(subset(flightfull, flightfull$markers == "F"))
# Summarize Just flight VCO2 data
overallflight <- c(summary(justflight$VCO2), sd=sd(justflight$VCO2, na.rm=T))
#mean, sd, and max
meanflightVCO2 <- overallflight[4]
sdflightVCO2 <- sd(justflight$VCO2, na.rm=T)
maxflightVCO2 <- overallflight[6]
```

```{r}
# longest continuous flight subset 
subnewend <- subset(justflight, justflight$markz != "same")
#View(subnewend) # tells all the start / ends of the behaviors
contigseconds <- data.frame(subnewend$Seconds[which(subnewend$markz == "new")], subnewend$Seconds[which(subnewend$markz == "end")])
names(contigseconds) <- c("new", "end")
contigseconds$length <- contigseconds$end - contigseconds$new
startseconds <- contigseconds$new[which(contigseconds$length == max(contigseconds$length))]
if(length(startseconds)  > 1) {startseconds <- startseconds[1]}
endseconds <- contigseconds$end[which(contigseconds$length == max(contigseconds$length))]
if(length(endseconds) > 1) {endseconds <- endseconds[1]}
######I get FALSE when I run this line#########################
60 < endseconds - startseconds #Make sure this is true, that the longest flight is greater than 60 seconds
######I get TRUE when I run this line#########################
30 < endseconds - startseconds

xlx <- ifelse((60 < endseconds - startseconds)==T, 60, (endseconds - startseconds))

z <- subset(justflight, justflight$Seconds >= startseconds)
useflight <- subset(z, z$Seconds <= endseconds) # This is longest continuous flight 
```

```{r}
##### Chosen Regions MR determination------
# 1. 
#     Overall contiguous  flight
(overallcontig <- c(summary(useflight$VCO2), sd=sd(useflight$VCO2, na.rm=T)))
(meancontig <- overallcontig[4])
(sdcontig <- sd(useflight$VCO2, na.rm=T))
(maxcontig <- overallcontig[6])
(lengthcontig <- length(useflight$Seconds))
#plot(useflight$VCO2 ~ useflight$Seconds, type="l")

# 2. 
#     Most stable 1 minute of flight in longest contiguous stretch
allVar <- data.frame(startTime=integer(0), time=integer(0), varVal=numeric(0))
for(x in 1:(length(useflight$Seconds)-xlx)) {
  varNow <- var(useflight$VCO2[x:(x+xlx)])
  timenow <- useflight$Seconds[x]
  newRow <- c(x, timenow, varNow)
  allVar <- rbind(allVar, newRow)  
}

names(allVar) <- c("startTime", "time", "varVal")
#View(allVar)
#now to find smallest variance
minVar <- min(allVar$varVal, na.rm=T)

(begin.sec <- allVar$startTime[which(allVar$varVal == minVar)])
duration.sec <- xlx
flightlowvar <- begin.sec:(begin.sec+duration.sec)
(MRlowvar <- c(summary(useflight$VCO2[begin.sec:(begin.sec+duration.sec)]), sd = sd(useflight$VCO2[begin.sec:(begin.sec+duration.sec)]), start=allVar$time[which(allVar$varVal == minVar)]))

(meanstable1min <- MRlowvar[4])
(sdstable1min <- sd(useflight$VCO2[begin.sec:(begin.sec+duration.sec)]))
(maxstable1min <- MRlowvar[6])

# 3. 
#     Highest mean 1 minute during contiguous flight
allhighmean <- data.frame(startTime=integer(0), timenow = integer(0), meanVal=numeric(0))

#This for loop computes the VCO2 mean for the next 1 minute for each timepoint 
for(x in 1:(length(useflight$Seconds)-xlx)){
  meanNow <- mean(useflight$VCO2[(x):(x+(xlx))], na.rm = T)
  timenow <- useflight$Seconds[x]
  newRow <- c(x, timenow, meanNow)
  allhighmean <- rbind(allhighmean, newRow)  
}
names(allhighmean)=c("startTime", "time", "meanVal")
#View(allhighmean)

#Get the estimates
beginsecmean <- allhighmean$startTime[which(allhighmean$meanVal == max(allhighmean$meanVal, na.rm = T))]
durationsecmean <- xlx
highmean.region <- beginsecmean:(beginsecmean+durationsecmean)
(summaryhighmean <- c(summary(useflight$VCO2[highmean.region], na.rm=T), sd=sd(useflight$VCO2[highmean.region], na.rm=T), start=beginsecmean, end=beginsecmean+durationsecmean))
(meanhighregion <- summaryhighmean[4])
(sdhighregion <- sd(useflight$VCO2[highmean.region], na.rm=T))
(maxhighregion<- summaryhighmean[6])

#4.   Highest 120 points during longest contiguous flight
n <- 60
decendingVCO2 <- sort(useflight$VCO2, decreasing = TRUE)
high120 <- decendingVCO2[1:n]
(regionhighest120 <- c(summary(high120), sd = sd(high120, na.rm=T)))
(meanhighest120 <- regionhighest120[4])
(sdhighest120 <- sd(high120, na.rm=T))
(maxhighest120 <- regionhighest120[6])

#plot(startflight$Seconds, startflight$VCO2, type="l", main = paste(filename, "(longest continuous)"))
#par(new=T)
#plot(startflight$Seconds[which(flightfull$markers == "F")], startflight$VCO2[which(flightfull$markers == "F")], col="blue", axes=T, ann=F)

#5.   Integral ... sort of .... 
#Sum  of all points in 
# The most stable 1 minute
#(intregralstable1min <- sum(useflight$VCO2[begin.sec:(begin.sec+60)]))
 
# Highest mean 1 minute
#(integralhigh1min <- sum(useflight$VCO2[beginsecmean:(beginsecmean+60)]))
```

```{r}
###### Exporting graphs and numbers to excel docs etc. ----

# Export graphs of whole trace + each determined region + peaks etc. 
analysis.file <- paste(filename, "traces.pdf", sep="-")
pdf(paste(dir.processed, analysis.file, sep="/"), width=10, height=7)


plot(flightfull$Seconds, flightfull$wvpnolag, type="l", col="blue", xlab = "Seconds", main = paste(filename, "- No Lag, uncorrected for drift"))
par(new=T)
plot(flightfull$Seconds, flightfull$co2nolag, type="l", col="darkorange", axes=F, ann=F)
par(new=T)
plot(flightfull$Seconds, flightfull$o2nolag, type="l", col="green", axes=F, ann=F)
par(new=T)
legend(x="topright", legend=c("WVP", "CO2", "O2"), col=c("blue", "darkorange","green"), bty = "n", lty=1, cex=.75)

plot(flightfull$Seconds, flightfull$VCO2, type="l", main = paste("VCO2", filename, "full trace"), xlab= "Seconds" , ylab = "VCO2 ml/hr")
abline(v=fly1)

plot(flightfull$Seconds[fly], flightfull$VCO2[fly], type="l", xlab= "Seconds" , ylab = "VCO2 ml/hr", main = paste("VCO2", filename, "during flight"))
abline(v=fly1)

plot(useflight$Seconds, useflight$VCO2, type="l", main = paste("VCO2 ", filename, " Longest contiguous"), ylab = "VCO2 ml/hr", xlab = "Seconds")

plot(useflight$Seconds[highmean.region], useflight$VCO2[highmean.region], main=c(filename, "VCO2 highest 1 minute"), type="l", xlab = "Seconds", ylab = "VCO2 ml/hr")

plot(useflight$Seconds[flightlowvar], useflight$VCO2[flightlowvar], type = "l", xlab=paste("Seconds (start:",begin.sec,"s)"), ylab= "VCO2 ml/hr", main = paste(filename, " Stable 1 min"))

dev.off()
```

```{r}
### EXPORT HARD MR NUMBERS

# All flight
#maxflightVCO2 <- overallflight[6]
#meanflightVCO2 <- overallflight[4]
#sdflightVCO2 <- sd(justflight$VCO2, na.rm=T)

# whole continguous 
#meancontig <- overallcontig[4]
#sdcontig <- overallcontig[7]
#maxcontig <- overallcontig[6]
#lengthcontig <- length(useflight$Seconds)

# stable 1 minutes
#(meanstable1min <- MRlowvar[4])
#(sdstable1min <- MRlowvar[7])
#(maxstable1min <- MRlowvar[6])

# highest mean 1 minute
#(meanhighregion <- summaryhighmean[4])
#(sdhighregion <- summaryhighmean[7])
#(maxhighregion<- summaryhighmean[6])

# highest 120 points
#meanhighest120 <- regionhighest120[4]
#sdhighest120 <- regionhighest120[7]
#maxhighest120 <- regionhighest120[6]



#Combine the metrics of RMR into one row and add to the a read in data file 
this_MR_data <- c(filename, meanflightVCO2, sdflightVCO2, maxflightVCO2, meancontig, sdcontig, maxcontig, lengthcontig, meanstable1min, sdstable1min, maxstable1min,
                  meanhighregion, sdhighregion, maxhighregion, meanhighest120, sdhighest120, maxhighest120)


FMR_all <- read.csv("FMR_all.csv", header = T, sep=",")
FMR_all$Filename <- as.character(FMR_all$Filename)
str(FMR_all)


#Add newest processed row to previous data, overwrite the csv (then move on to the next one)
FMR_all <- rbind(FMR_all, this_MR_data)
write.csv(FMR_all, "FMR_all.csv", row.names=F)



prelowvar
lengthcontig


# check then clear the workspace
#ls()
rm(list = ls())



```


